quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-07-07 paper-quollr.Rmd
Replaced the word visualisation

The paper is organized as follows. The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

### Visualizations

The package offers several $2-\text{D}$ visualizations, including:

@@

Each visualization can be generated using its respective function, as described in this section.

@@

#### $p\text{-}D$ model visualization

To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualization using the `show_langevitour()` function. This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.

@@

The `show_langevitour()` function then renders the visualization using the `langevitour` interface, displaying both types of points in a dynamic tour. The `edge_data` input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model’s structure.

In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.

@@

To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`. This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualization, the input data must be prepared using the `comb_all_data_model()` function. This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2-\text{D}$ bin layout (`model_2d`):

@@

Before visualization, the input data must be prepared using the `comb_all_data_model_error()` function. The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2-\text{D}$ model data (`model_2d`), and model error (`error_data`).

>>>>

The paper is organized as follows. The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualisation capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

### visualisations

The package offers several $2-\text{D}$ visualisations, including:

@@

Each visualisation can be generated using its respective function, as described in this section.

@@

#### $p\text{-}D$ model visualisation

To evaluate how well the $p\text{-}D$ model captures the underlying structure of the high-dimensional data, we provide a visualisation using the `show_langevitour()` function. This function renders a dynamic projection of both the high-dimensional data and the model using the `langevitour` package.

@@

The `show_langevitour()` function then renders the visualisation using the `langevitour` interface, displaying both types of points in a dynamic tour. The `edge_data` input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model’s structure.

In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the $2-\text{D}$ wireframe structure mapped back to high-dimensional space.

@@

To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`. This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualisation, the input data must be prepared using the `comb_all_data_model()` function. This function combines the high-dimensional data (`highd_data`), the NLDR data (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the $2-\text{D}$ bin layout (`model_2d`):

@@

Before visualisation, the input data must be prepared using the `comb_all_data_model_error()` function. The function requires several arguments: points data which contain high-dimensional data (`highd_data`), NLDR data (`nldr_data`), high-dimensional model data (`model_highd`), $2-\text{D}$ model data (`model_2d`), and model error (`error_data`).

### 2025-07-07 paper-quollr.tex
Replaced the word visualisation

The paper is organized as follows. The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

\subsubsection{Visualizations}\label{visualizations}

The package offers several \(2-\text{D}\) visualizations, including:

@@

\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}

To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualization using the \texttt{show\_langevitour()} function. This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.

@@

The \texttt{show\_langevitour()} function then renders the visualization using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour. The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.

In the resulting interactive visualization black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.

@@

To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}. This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function. This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):

@@

Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function. The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).

>>>>

The paper is organized as follows. The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualisation capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

\subsubsection{visualisations}\label{visualisations}

The package offers several \(2-\text{D}\) visualisations, including:

@@

\paragraph{\texorpdfstring{\(p\text{-}D\) model visualisation}{p\textbackslash text\{-\}D model visualisation}}\label{ptext-d-model-visualisation}

To evaluate how well the \(p\text{-}D\) model captures the underlying structure of the high-dimensional data, we provide a visualisation using the \texttt{show\_langevitour()} function. This function renders a dynamic projection of both the high-dimensional data and the model using the \texttt{langevitour} package.

@@

The \texttt{show\_langevitour()} function then renders the visualisation using the \texttt{langevitour} interface, displaying both types of points in a dynamic tour. The \texttt{edge\_data} input defines connections between neighboring bins (i.e., the hexagonal edges) to visualize the model's structure.

In the resulting interactive visualisation black points represent the high-dimensional data, green points represent the model centroids from each bin, and the lines between model points reflect the \(2-\text{D}\) wireframe structure mapped back to high-dimensional space.

@@

To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}. This visualisation combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function. This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR data (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the \(2-\text{D}\) bin layout (\texttt{model\_2d}):

@@

Before visualisation, the input data must be prepared using the \texttt{comb\_all\_data\_model\_error()} function. The function requires several arguments: points data which contain high-dimensional data (\texttt{highd\_data}), NLDR data (\texttt{nldr\_data}), high-dimensional model data (\texttt{model\_highd}), \(2-\text{D}\) model data (\texttt{model\_2d}), and model error (\texttt{error\_data}).
